Exploring the Feasibility of\\ Multifactor Authentication with EarEEG

### 2018-02-06 paper.tex
Minor fixes

\title{Exploring the Feasability of\\ Multifactor Authentication with EarEEG}

@@

In the proposed system we seek to incorporate the perspectives of this research. Passthoughts such as relaxed breathing or imagining favorite song are easy for a user to remember and perform, and the earpiece form factor of our system is intended to address wearability and social perception concerns.

@@

Seven male, right-handed participants (P1-P7), five students and two non-students, completed our study protocol approved by our local ethics review board. Though this sample is relatively homogenous and greater diversity is necessary for a larger real-world feasibility assessment, this quality interestingly functions to strengthen the results of this initial exploration (see details in Discussion). After participants' 3D ear molds were obtained and their fit and electrical impedances were checked, the collection of data for authentication analyses commenced. Data collection consisted of participants completing a demographics questionnaire, a set up period with the OpenBCI system and earpieces with a second impedance check, their performance of nine mental tasks, and finally a post-experiment questionnaire.

@@

Since past work has shown that classification tasks in EEG-based brain-computer interfaces (BCI) are linear \cite{Garrett2003a}, we used XGBoost, a popular tool for logistic linear classification \cite{Chen2016}, to analyze the mental task EEG data. Compared to other linear classifiers, XGBoost uses gradient boosting in which an algorithm generates aa decision tree of weak linear classifiers that minimizes a given loss function. Gradient boosting generally improves linear classification results without manually tuning hyperparameters.

@@

Before the end of othe session, participants completed a usability questionnaire. This questionnaire asked them to rate each mental task on four 7-point likert-type scales: ease of use, level of engagement, perceived repeatability, and likeliness to use in a real-world authentication setting. The \textit{breathe} and \textit{listen} tasks were rated easiest to use (\(\mu\)=6.75), \textit{Sequence}, \textit{song}, and \textit{song - open} as most engaging (\(\mu\)=5), \textit{Breathe} (\(\mu\)=7) and \textit{listen} (\(\mu\)=6.75) as most repeatable. For likeliness to use for authentication, \textit{song - open} (\(\mu\)=5) and \textit{sequence} (\(\mu\)=4.25) were rated highest, though modestly. Participants also ranked the tasks overall from most (1) to least (9) favorite. \textit{Song - open} ranked most favorites (\(\mu\)=4.25) followed by a tie between \textit{breathe - open}, \textit{song}, and \textit{speech} (\(\mu\)=4.75). \textit{Sequence} (\(\mu\)=7.75) and \textit{face} (\(\mu\)=6.75) were ranked least favorite.

@@

Our multiple classifier training strategies establish the presence of both knowledge and inherence factors as schemes isolating each still achieve low FARs. Interestingly, by training classifiers to reject both wrong participants and wrong tasks from the correct participant, thus incorporating both the knowledge and inherence factors, we see an increase in FRR suggesting multi-factor authentication's added security may come at a small cost to convenience. As preliminary evidence of the potential \textit{secondary} inherence factor of ear shape in the current prototypes, we tested P1 using their best task (listen) with P6's earpieces but tested on P1's classifier, with a resultant FAR of 0, suggesting the protective value of the custom-fit earpieces likely due to differing electrical impedances from one user to another. To add a possession factor, these custom-fit earpieces could also easily include a hardware keypair to sign authentication attempts, functioning similarly to the way a key fob can be used to unlock a car, but compared to this single-factor system the demonstrated inherence and knowledge factors would provide important extra layers of security.

@@

As demonstrated by these preliminary results, custom-fit, in-ear EEG earpieces can provide three factors of security in one highly usable step: thinking one's passthought, using the discreet form factor of an earpiece. Among this initial sample, we are able to achieve 99.82\% authentication accuracy using a single sensing earpiece, showing potential for integration with technology already used in everyday life (like earphones). By expanding our corpus of EEG readings (in population size, time, and diversity of settings), we hope to better understand the underlying distribution of EEG signals and security properties of passthoughts as well usability issues that may arise in different contexts.

>>>>

\title{Exploring the Feasibility of\\ Multifactor Authentication with EarEEG}

@@

In the proposed system we seek to incorporate the perspectives of this research. Passthoughts such as relaxed breathing or imagining a favorite song are easy for a user to remember and perform, and the earpiece form factor of our system is intended to address wearability and social perception concerns.

@@

Seven male, right-handed participants (P1-P7), five students and two researchers, were recruited via a university mailing list and completed our study protocol approved by our local ethics review board. Though this sample is relatively homogenous and greater diversity is necessary for a larger real-world feasibility assessment beyond this visionary exploration, this quality interestingly functions to strengthen the results of a system designed to discriminate between users (see Discussion). After participants' 3D ear molds were obtained and their fit and electrical impedances were checked, the collection of data for authentication analyses commenced. Data collection consisted of participants completing a demographics questionnaire, a set up period with the OpenBCI system and earpieces with a second impedance check, their performance of nine mental tasks, and finally a post-experiment questionnaire.

@@

Since past work has shown that classification tasks in EEG-based brain-computer interfaces (BCI) are linear \cite{Garrett2003a}, we used XGBoost, a popular tool for logistic linear classification \cite{Chen2016}, to analyze the mental task EEG data. Compared to other linear classifiers, XGBoost uses gradient boosting in which an algorithm generates a decision tree of weak linear classifiers that minimizes a given loss function. Gradient boosting generally improves linear classification results without manually tuning hyperparameters.

@@

Before the end of the session, participants completed a usability questionnaire. This questionnaire asked them to rate each mental task on four 7-point likert-type scales: ease of use, level of engagement, perceived repeatability, and likeliness to use in a real-world authentication setting. The \textit{breathe} and \textit{listen} tasks were rated easiest to use (\(\mu\)=6.75), \textit{Sequence}, \textit{song}, and \textit{song - open} as most engaging (\(\mu\)=5), \textit{Breathe} (\(\mu\)=7) and \textit{listen} (\(\mu\)=6.75) as most repeatable. For likeliness to use for authentication, \textit{song - open} (\(\mu\)=5) and \textit{sequence} (\(\mu\)=4.25) were rated highest, though modestly. Participants also ranked the tasks overall from most (1) to least (9) favorite. \textit{Song - open} ranked most favorites (\(\mu\)=4.25) followed by a tie between \textit{breathe - open}, \textit{song}, and \textit{speech} (\(\mu\)=4.75). \textit{Sequence} (\(\mu\)=7.75) and \textit{face} (\(\mu\)=6.75) were ranked least favorite.

@@

Our multiple classifier training strategies establish the presence of both knowledge and inherence factors as schemes isolating each still achieve low FARs. Interestingly, by training classifiers to reject both wrong participants and wrong tasks from the correct participant, thus incorporating both the knowledge and inherence factors, we see an increase in FRR suggesting multi-factor authentication's added security may come at a small cost to convenience. As preliminary evidence of the potential \textit{secondary} inherence factor of ear shape in the current prototypes, we tested P1 using their best task (listen) with P6's earpieces but tested on P1's classifier, with a resultant FAR of 0, suggesting the protective value of the custom-fit earpieces likely due to differing electrical impedances from one user to another. To add a possession factor, these custom-fit earpieces could also easily include a hardware keypair to sign authentication attempts, functioning similarly to the way an electronic key fob can be used to unlock a car, but compared to this single-factor system the demonstrated inherence and knowledge factors would provide important extra layers of security.

@@

As demonstrated by these preliminary results, custom-fit, in-ear EEG earpieces can provide three factors of security in one highly usable step: thinking one's passthought wearing a discreet earpiece. Among this initial sample, we are able to achieve 99.82\% authentication accuracy using a single sensing earpiece, showing potential for integration with technology already used in everyday life (like earphones). By expanding our corpus of EEG readings (in population size, time, and diversity of settings), we hope to better understand the underlying distribution of EEG signals and security properties of passthoughts as well usability issues that may arise in different contexts.
